---
title: "cma-week5"
author: "Martine Besse"
date: "2023-05-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("readr")
library("sf")
library("terra")
library("dplyr")
library("lubridate")
library("ggplot2")
library("tmap")
options("sf_max.plot" = 1) #so base plotting only take 1

wildschwein_BE <- read_delim("datasets/wildschwein_BE_2056.csv", ",") |>
  st_as_sf(coords = c("E", "N"), crs = 2056, remove = FALSE)
```

#Task 1
note: ggplot use by default WGS84 for the axis  
```{r}
#equivalent: read_sf (quieter) / st_read
fanel <- st_read("datasets/Feldaufnahmen_Fanel.gpkg")
#plot(fanel["Frucht"])


ggplot(fanel) +
  geom_sf(aes(fill = Frucht)) +
  geom_sf(data = wildschwein_BE, alpha = .3, pch = ".") + #pch = point type
  theme(legend.position = "none") + # to remove the legend
  coord_sf(datum = 2056)
```

spatial join: we want to link wildboard to the field.  
annotate wildboar data with info on Land use
```{r}
wildschwein_BE <- st_join(wildschwein_BE, fanel) #|> View()

ggplot(fanel) +
  #geom_sf(aes(fill = Frucht, alpha = .3)) +
  geom_sf(data = wildschwein_BE, aes(colour = Frucht), alpha = .3, pch = ".") + #pch = point type
  theme(legend.position = "none") + # to remove the legend
  coord_sf(datum = 2056)


```

#Task4
Raster:  
class       : SpatRaster 
dimensions  : 5303, 7001, 1  (nrow, ncol, nlyr)
resolution  : 1, 1  (x, y)
extent      : 2568153, 2575154, 1202306, 1207609  (xmin, xmax, ymin, ymax)
coord. ref. : +proj=somerc +lat_0=46.9524055555556 +lon_0=7.43958333333333 +k_0=1 +x_0=2600000 +y_0=1200000 +ellps=bessel +units=m +no_defs 
source      : vegetationshoehe_LFI.tif 
name        : vegetationshoehe_LFI 
min value   :                 0.00 
max value   :                47.58 

nlyr = number of layers  

```{r}
veg <- rast("datasets/vegetationshoehe_LFI.tif") 
#plot(veg)

tm_shape(veg) +
  tm_raster(style = "cont", palette = "viridis") + #style continuous
  tm_shape(wildschwein_BE) + tm_dots()

#extracting the values of the raster (vegetation height) for each point
#we can use attributes to specify how summarize the extracted data "FUN"
veg_extract <- extract(veg, wildschwein_BE)
str(veg_extract)

wildschwein_BE$veg <- veg_extract$vegetationshoehe_LFI
wildschwein_BE


ggplot(fanel) +
  #geom_sf(aes(fill = Frucht, alpha = .3)) +
  geom_sf(data = wildschwein_BE, aes(colour = veg), alpha = .3, pch = ".") + #pch = point type
  theme(legend.position = "none") + # to remove the legend
  coord_sf(datum = 2056)
```

'data.frame':	51246 obs. of  2 variables:
 $ ID                  : num  1 2 3 4 5 6 7 8 9 10 ...
 $ vegetationshoehe_LFI: num  20.1 23.9 25 21.6 15.7 ...
 ->  51246 = n° of features in the wildboar data
 
 
 
 Example with buffered point  
 Because there is a certain level of uncertainty around GPS points.  
```{r}
wildschwein_buffer <- st_buffer(wildschwein_BE, 20)
wildschwein_buffer
plot(wildschwein_buffer)

veg_extract2 <- extract(veg, wildschein_buffer, fun = mean)

```
 
